\documentclass{oxmathproblems}

\printanswers

\course{Understanding Analysis, 2nd Edition: Stephen Abbott}
\sheettitle{Chapter 3 - Section 3.2 - Open and Closed Sets} %can l7ave out if no title per sheet

\begin{document}
\begin{questions}

%------------------------- Problem 1 -------------------------
\miquestion
\begin{enumerate}[label=(\alph*)]
  \item Where in the proof of Theorem 3.2.3 part (ii) does the assumption that the collection of open sets be \textit{finite} get used?
  \item Give an example of a countable collection of open sets $\{O_{1}, O_{2}, O_{3}, ...\}$ whose intersection $\cap_{n=1}^{\infty}O_{n}$ is closed, not empty, and
  not all of $R$.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item The part where we want to take the minimum \\
    $\epsilon$-neighborhood $\epsilon = min\{\epsilon_{1}, \epsilon_{2}, ..., \epsilon_{n}\} > 0$. A minimum does not necessarily exist in an infinite set.

    To be even more precise though, what we're lacking is a positive lower bound. Had there been a positive infimum (but not minimum), we could use that as $\epsilon$ (note
    that the proof is still valid even if we replace minimum with infimum. It's just that taking minimum is simpler for a finite set). However, the infimum may
    be zero, as can be seen in the next example.

    \item Let $O_{n} = (-\frac{1}{n}, \frac{1}{n})$. Both endpoints converge to 0, so $\cap_{n=1}^{\infty}O_{n} = \{0\}$ which is closed, not empty, and not all of $R$, as required.
  \end{enumerate}
\end{solution}

%------------------------- Problem 2 -------------------------
\miquestion Let $A = \{(-1)^{n} + \frac{2}{n} : n = 1, 2, 3, ...\}$ and $B = \{x \in Q : 0 < x < 1\}$. Answer the following questions for each set:
\begin{enumerate}[label=(\alph*)]
  \item What are the limit points?
  \item Is the set open? Closed?
  \item Does the set contain any isolated points?
  \item Find the closure of the set.
\end{enumerate}

%------------------------- Problem 3 -------------------------
\miquestion Decide whether the following sets are open, closed, or neither. If a set is not open, find a point in the set for which there is $\epsilon$-neighborhood contained in
the set. If a set is not closed, find a limit point that is not contained in the set.
\begin{enumerate}[label=(\alph*)]
  \item \textbf{Q}.
  \item \textbf{N}.
  \item \{$x \in \textbf{R}$ : $x \neq 0$\}.
  \item \{$1 + 1/4 + 1/9 + ... + 1/n^{2} : n \in \textbf{N}$\}.
  \item \{$1 + 1/2 + 1/3 + ... + 1/n : n \in \textbf{N}$\}.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item Neither. 
    
    To prove it's not open, take any rational $r$ and any interval $(r-\epsilon, r+\epsilon)$. If \textbf{Q} is open then every number in the interval must be in
    \textbf{Q}. But between any two real numbers there is always an irrational.

    To prove it's not closed, we use the sequence from Heron's method (Exercise 2.4.5). Let $x_{1} = 2$ and define the sequence $x_{n+1} = \frac{1}{2}(x_{n} + \frac{2}{x_{n}})$.
    Then each $x_{n}$ is rational and $(x_{n})$ converges to $\sqrt{2}$ which is not in \textbf{Q}.
  
    \item It's not open. Take any $n \in \textbf{N}$, any interval $(n-\epsilon, n+\epsilon)$ must contain a non-integer.
    
    But it is closed. A convergent sequence must be bounded. If a sequence in \textbf{N} is bounded then the set of numbers in the sequence is finite.
    But for it to be convergent, the only way is for a number $n \in \textbf{N}$ to eventually repeat* and be the limit. Since $n \in \textbf{N}$, $\textbf{N}$ is closed.

    *a rigorous proof of this statement. If the sequence is convergent, then by Cauchy Criterion there exists $N \in \textbf{N}$ such that for all $m, n \geq N$, $|a_{n} - a_{m}| < \epsilon$
    Take $\epsilon < 1$. Since the natural numbers are at least 1 apart from each other, the only way to satisfy the inequality is for $a_{n} = a_{m}$.
  
    \item Open (e.g. take $\epsilon = \frac{|x|}{2}$ for any $x \neq 0$), but not closed. The sequence $(\frac{1}{n^2})$ is in the set but its limit 0 is not.
    
    \item Not open. 
    
    (Not quite correct) Let $x_{n} = \sum_{k=1}^{n}\frac{1}{k^2}$. Given any $\epsilon > 0$, the neighborhood $(x_{n}-\epsilon, x_{n}+\epsilon)$ does not contain $\frac{x_{n} + x_{n+1}}{2}$.

    The proof above is only valid if the midpoint is actually inside $(x_{n}-\epsilon, x_{n}+\epsilon)$. To prove not open, we need to produce, for \textit{any} $\epsilon$, a
    point that's in the neighborhood but not in the set. If $\epsilon$ is small enough, the midpoint may not be in the neighborhood at all.

    (Corrected proof). To ensure that the missing point is always in the neighborhood, we can instead take min($x_{n}+\frac{\epsilon}{2}$, $\frac{x_{n} + x_{n+1}}{2}$). We take
    $\epsilon/2$ instead of $\epsilon$ because $x+\epsilon$ is outside the neighborhood (it's right on the boundary).
    
    Not closed either, as $\sum_{k=1}^{\infty}\frac{1}{k^2}$ converges to a limit $L \in \textbf{R}$ (as per Example 2.4.4), but $L$ is not in the set, because any element in the
    set is equal a \textit{finite} partial sum.

    \item Not open. Like the previous part, the point min($x_{n}+\frac{\epsilon}{2}$, $\frac{x_{n} + x_{n+1}}{2}$) is in the neighborhood but not in the set.
    
    Closed, for a similar reason as part (b): for a sequence in this set to converge, an element must eventually repeat. But proving it with Cauchy Criterion is tricky this time,
    as the distance between the elements is not constant.

    Instead, we can use subsequence convergence. Let $(x_{n})$ be a convergent sequence with elements from the set. $(x_{n})$ must only contain a finite number of distinct terms,
    for otherwise $\sum_{k=1}^{\infty}\frac{1}{n}$, which diverges, would make $(x_{n})$ unbounded.
    
    Then an element $x_{k}$ must appear infinitely many times in $(x_{n})$. Take the
    subsequence of all $x_{k}$'s. Since $(x_{n})$ converges, any of its subsequence must converge to the same limit. The subsequence $(x_{k})$ converges to $x_{k}$ which is in the set.
  \end{enumerate}
\end{solution}

%------------------------- Problem 4 -------------------------
\miquestion Let $A$ be nonempty and bounded above so that $s = \text{sup } A$ exists.
\begin{enumerate}[label=(\alph*)]
  \item Show that $s \in \overline{A}$.
  \item Can an open set contain its supremum?
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item We only need to show that $s$ is a limit point of $A$. Then by the definition of closure, $s \in \overline{A}$.
    
    $s$ being a limit point follows from the definition of supremum. We need to prove that there is a sequence in $A$ that converges to $s$ i.e. for all $\epsilon > 0$,
    there is an $x \in A$ with $|s-x| < \epsilon$. Suppose for contradiction that such an $x$ doesn't exist, then $s-\epsilon$ is a lower bound of $A$ that's less
    than the supremum, contradicting the fact that $s$ is the \textit{least} upper bound.

    \item No. Take any $\epsilon$-neighborhood of the supremum, then by defintion of supremum the interval $(s, s+\epsilon)$ doesn't contain any element of $A$.
  \end{enumerate}
\end{solution}

%------------------------- Problem 5 -------------------------
\miquestion Prove Theorem 3.2.8. A set $F \subseteq \textbf{R}$ is closed if and only if every Cauchy sequence contained in $F$ has a limit that is also an element of $F$.
\begin{solution}
  The only thing we need is Theorem 2.6.4 (Cauchy Criterion): A sequence converges if and only if it's Cauchy.

  To prove the "only if" part, suppose $F$ is closed. Then $F$ contains its limit points. Then every convergent sequence in $F$ has a limit that's in $F$. Since every Cauchy sequence
  is convergent, we're done. (It's also worth noting that the limit of the \textit{sequence} may be a limit \textit{point} of the \textit{set} or an isolated point but in either case, it's in $F$).

  To prove the other direction, suppose every Cauchy sequence has a limit in $F$. Since every convergent sequence is Cauchy, then every convergent sequence also has a limit in $F$. Then
  $F$ is closed by definition.
\end{solution}

%------------------------- Problem 6 -------------------------
\miquestion Decide whether the following statements are true or false. Supply proofs or counterexample.
\begin{enumerate}[label=(\alph*)]
  \item An open set that contains every rational number must necessarily be all of \textbf{R}.
  \item The Nested Interval Property remains true if the term "closed interval" is replaced by "closed set".
  \item Every nonempty open set contains a rational number.
  \item Every bounded infinite closed set contains a rational number.
  \item The Cantor Set is closed.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item False. Consider the set $\textbf{R} \setminus \{\sqrt{2}\}$. Any point $x$ has a neighborhood $(x-\epsilon, x+\epsilon)$ with $\epsilon = \frac{|x-\sqrt{2}|}{2}$, and
    all numbers in the neighborhood are in the set.

    \item False. Let $X_{1} \supseteq X_{2} \supseteq X_{3} \supseteq ...$ with $X_{n} = [n, \infty)$. The intersection is empty because
    any number $x$ is not in $X_{\lceil x \rceil + 1}$.

    As will be seen in the next section, for this to be true it requires compactness i.e. closed \textit{and} bounded.
    
    \item True. Take any point $x$ from the set. Since it's open, there is a neighborhood $(x-\epsilon, x+\epsilon)$ where all numbers are in the set. By density of the rationals,
    this neighborhood must contain a rational number.

    \item False. Consider the set $\{\sqrt{2}, \frac{\sqrt{2}}{1+1}, \frac{\sqrt{2}}{1+\frac{1}{2}}, \frac{\sqrt{2}}{1+\frac{1}{3}}, \frac{\sqrt{2}}{1+\frac{1}{4}}, ...\}$.
    It's bounded by $\frac{\sqrt{2}}{2}$ and $\sqrt{2}$, infinite, and closed, but contains no rationals.

    \item True. As per page 85, the Cantor Set can be defined as $\bigcap_{n=0}^{\infty} C_{n}$ where each $C_{n}$ is a finite union of closed intervals. By Theorem 3.2.14(i),
    the union of a finite collection of closed sets is closed, so each $C_{n}$ is closed. By the part (ii) of the same theorem, the intersection of an arbitrary collection of
    closed sets is closed.
  \end{enumerate}
\end{solution}

%------------------------- Problem 7 -------------------------
\miquestion Given $A \subseteq \textbf{R}$, let $L$ be the set of all limit points of $A$.
\begin{enumerate}[label=(\alph*)]
  \item Show that the set $L$ is closed.
  \item Argue that if $x$ is a limit point of $A \cup L$, then $x$ is a limit point of $A$. Use this observation to furnish a proof for Theorem 3.2.12 (for any $A \subseteq \textbf{R}$,
  the closure $\overline{A}$ is a closed set and is the smallest set containing $A$).
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item Suppose $X$ is a limit point of $L$. If $L$ is closed, then $X \in L$ is also a limit point of $A$. That gives an idea of the proof: there's a sequence in $A$ that converges to $X$.
    
    $X$ is a limit point, so there's a sequence $(l_{n})$ that converges to $X$. For any $\epsilon > 0$, there exists $N$ such that for all $n \geq N$, $|l_{n} - X| < \epsilon$. The idea
    is that since each $l_{n}$ is a limit point of $A$, we can take an $a \in A$ arbitrarily close to $l_{n}$, which would make it close to $X$ as well.

    Concretely, construct $(a_{n})$ such that $|a_{n} - l_{n}| < 1/n$. Now, there's an $N_{1}$ such that $n \geq N_{1} \implies |l_{n} - X| < \epsilon/2$. There's also an $N_{2}$
    such that $n \geq N_{2} \implies |a_{n} - l_{n}| < \epsilon/2$. Then $|a_{n} - X| \leq |a_{n} - l_{n}| + |l_{n} - X| < \epsilon/2 + \epsilon/2 = \epsilon$.

    \item To prove $x$ is a limit point of $A$, use a very similar argument as part (a). I'm not really in the mood for writing it down right now. Maybe some other day\dots
    
    Once that's done, we go back to the proof of Theorem 3.2.12 in the book. It's missing the part where we need to prove that $A \cup L$ does not contain any "new" limit points
    outside of $L$. But we've proved that the limit points of $A \cup L$ are also limit points of $A$, all of which are in $L$.
  \end{enumerate}

\end{solution}

\end{questions}
\end{document}
