\documentclass{oxmathproblems}

\printanswers

\course{Understanding Analysis, 2nd Edition, Stephen Abbott}
\sheettitle{Chapter 4 - Section 4.3 - Continuous Functions} 

\begin{document}
\begin{questions}

%------------------------- Problem 1 -------------------------
\miquestion Let $g(x) = \sqrt[3]{x}$.
\begin{enumerate}[label=(\alph*)]
  \item Prove that $g$ is continuous at $c = 0$.
  \item Prove that $g$ is continuous at a point $c \neq 0$. (The identity $a^3-b^3=(a-b)(a^2+ab+b^2)$ will be helpful.)
\end{enumerate}

\begin{solution} 
  \begin{enumerate}[label=(\alph*)]
    \item Pick $\delta = \epsilon^3$ then $|x| < \delta \implies |\sqrt[3]{x} - 0| < \sqrt[3]{\epsilon^3} = \epsilon$.
    \item $|\sqrt[3]{x} - \sqrt[3]{c}| = \frac{|x-c|}{|\sqrt[3]{x^2} + \sqrt[3]{xc} + \sqrt[3]{c^2}|}$. Since the denom must be positive, then
    \[\frac{|x-c|}{|\sqrt[3]{x^2} + \sqrt[3]{xc} + \sqrt[3]{c^2}|} \leq |x-c|\] Pick $\delta = \epsilon$.

    ...wait no. If the denom is less than 1, then the nume will get bigger. 
    
    Instead, since $c \neq 0$, we can always choose $\delta$ small enough so that $x$ and $c$ have the same sign (say, $\delta \leq \frac{|c|}{2}$), which would make all of the terms in the denom positive. Then:
    \[\frac{|x-c|}{|\sqrt[3]{x^2} + \sqrt[3]{xc} + \sqrt[3]{c^2}|} = \frac{|x-c|}{\sqrt[3]{x^2} + \sqrt[3]{xc} + \sqrt[3]{c^2}} \leq \frac{|x-c|}{\sqrt[3]{c^2}}\]
    Pick $\delta = \min(\frac{|c|}{2}, \epsilon\sqrt[3]{c^2})$.

    Another way is to note that $\sqrt[3]{x^2} + \sqrt[3]{xc} + \sqrt[3]{c^2}$ is a quadratic that's always positive with minimum value $\frac{3}{4}\sqrt[3]{c^2}$.
    Pick $\delta=\frac{3}{4}\epsilon\sqrt[3]{c^2}$.
  \end{enumerate}
\end{solution}

%------------------------- Problem 2 -------------------------
\miquestion In all of these, let $f$ be a function defined on all of $\textbf{R}$.
\begin{enumerate}[label=(\alph*)]
  \item Let's say that $f$ is \textit{onetinuous} at $c$ if for all $\epsilon > 0$ we can choose $\delta = 1$ and it follows that $|f(x)-f(c)| < \epsilon$
  whenever $|x-c| < \delta$. Find an example of a function that is onetinuous on all of $\textbf{R}$.
  \item Let's say that $f$ is \textit{equaltinuous} at $c$ if for all $\epsilon > 0$ we can choose $\delta = \epsilon$ and it follows that $|f(x)-f(c)| < \epsilon$
  whenever $|x-c| < \delta$. Find an example of a function that is equaltinuous on \textbf{R} that is nowhere onetinuous, or explain why there is no such function.
  \item Let's say that $f$ is \textit{lesstinuous} at $c$ if for all $\epsilon > 0$ we can choose $0 < \delta < \epsilon$ and it follows that $|f(x)-f(c)| < \epsilon$
  whenever $|x-c| < \delta$. Find an example of a function that is lesstinuous on \textbf{R} that is nowhere equaltinuous, or explain why there is no such function.
  \item Is every lesstinuous function continuous? Is every continuous function lesstinuous? Explain.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item The function $f(x) = k$ for a fixed constant $k$ is onetinuous.
    \item $f(x) = x$. It's equaltinuous but not onetinuous as given $\epsilon < 1$, $\delta = 1$ doesn't work. For example, let $\epsilon=\frac{1}{2}$ and
    $x = c+\frac{1}{2}$, then $|x-c| < 1$ but $|f(x)-f(c)| = \frac{1}{2}$ is not less than $\epsilon$.
    \item Intuitively, for $\delta < \epsilon$ to work, we need a function that grows faster than $f(x)=x$. 
    
    $f(x) = 3x$ works. Solve the $\epsilon-\delta$ to get $\delta < \frac{\epsilon}{3}$. Since $\delta < \frac{\epsilon}{3} < \epsilon$, the function is lesstinuous.

    It's nowehere equaltinuous. Let $x = c+\frac{\epsilon}{2}$. If $\delta=\epsilon$ then $(|x-c| = \frac{\epsilon}{2}) < \delta$, but $|f(x)-f(c)| = \frac{3\epsilon}{2} > \epsilon$.
  
    $f(x)=x^2$ also works. It's lesstinuous everywhere (too lazy to type the $\epsilon-\delta$ derivation). But it's not nowhere equaltinuous. Again choose
    $x=c+\frac{\epsilon}{2}$. Then $|f(x)-f(c)| = \epsilon|c| + \frac{\epsilon^2}{4}$. If this were equaltinuous, then for \textit{all} $\epsilon > 0$
    we need $|f(x)-f(c)| < \epsilon$. But for any $\epsilon > 4$ this is false.

    \item Yes to both.
    
    Every lesstinuous function is continuous. For all $\epsilon > 0$ there's a $\delta$ that works, and that's enough for continuity. It doesn't matter what the bound on $\delta$
    is, whether is lesstinuous or equaltinuous or whatever.

    Every continuous function is lesstinuous. By definition of continuity, for all $\epsilon > 0$ there's a $\delta$ such that $|x-c| < \delta \implies |f(x)-f(c)| < \epsilon$.
    If $\delta \geq \epsilon$, then a $\delta_{2} < \epsilon$ must also work, as it satisfies $|x-c| < \delta_{2} < \epsilon \leq \delta$.
  \end{enumerate}
\end{solution}

%------------------------- Problem 3 -------------------------
\miquestion
\begin{enumerate}[label=(\alph*)]
  \item Supply a proof for Theorem 4.3.9 using the $\epsilon-\delta$ characterization of continuity.
  \item Give another proof using the sequential characterization of continuity (from Theorem 4.3.2 (iii)).
\end{enumerate}

\begin{solution}
  To restate the (abridged) Theorem: we're given $f$ continuous at $c$ and $g$ continuous at $f(c)$ and we need to prove $g(f(x))$ is continuous at $c$.
  \begin{enumerate}[label=(\alph*)]
    \item To prove $g(f(x))$ continuous at $c$, then we need to prove that for all $\epsilon > 0$ there exists $\delta$ such that $|x-c| < \delta \implies |g(f(x)) - g(f(c))| < \epsilon$.
    
    From continuity of $f$ at $c$, then for all $\epsilon_{f} > 0$ there exists $\delta_{f}$ such that $|x-c| < \delta_{f} \implies |f(x)-f(c)| < \epsilon_{f}$.

    From continuity of $g$ at $f(c)$, then for all $\epsilon > 0$ there exists $\delta_{g}$ such that $|y-f(c)| < \delta_{g} \implies |g(y)-g(f(c))| < \epsilon$.

    Set $\epsilon_{f} = \delta_{g}$. Plug $f(x)$ as $y$, we have $|x-c| < \delta_{f} \implies |f(x)-f(c)| < \delta_{g} \implies |g(f(x))-g(f(c))| < \epsilon$, as required.

    Just don't get confused by the interchange between $y$ and $f(x)$. $y$ is just a placeholder name that stands for any input of $g$, of which the values of $f(x)$ is a part of.
 
    \item We need to prove that $(x_{n}) \to c \implies g(f(x_{n})) \to g(f(c))$.
    
    From continuity of $f$ at $c$, we have $(x_{n}) \to c \implies f(x_{n}) \to f(c)$.

    From continuity of $g$ at $f(c)$, we have $(y_{n}) \to f(c) \implies g(y_{n}) \to g(f(c))$.

    Plug $f(x_{n})$ as $(y_{n})$, then we have $(x_{n}) \to c \implies f(x_{n}) \to f(c) \implies g(f(x_{n})) \to g(f(c))$, as required.
  \end{enumerate}
\end{solution}

%------------------------- Problem 4 -------------------------
\miquestion Assume $f$ and $g$ are defined on all of \textbf{R} and that $\lim_{x \to p}f(x)=q$ and $\lim_{x \to q}g(x)=r$.
\begin{enumerate}[label=(\alph*)]
  \item Give an example to show that it may not be true that
  \[\lim_{x \to p}g(f(x)) = r.\]
  \item Show that the result in (a) does follow if we assume $f$ and $g$ are continuous.
  \item Does the result in (a) hold if we only assume $f$ is continuous? How about if we only assume that $g$ is continuous?
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
      \item Let $f(x) = 0$ everywhere and
      \[g(x) = 
        \begin{cases}
        0 & \text{if } x \ne 0 \\
        5 & \text{if } x = 0
        \end{cases}
      \]
      
      Set $p=0$. We have $\lim_{x \to 0}f(x) = 0$ and $\lim_{x \to 0}g(x) = 0$, but $\lim_{x \to 0}g(f(x))=g(0)=5$.
    \item See the next part. Continuity on $g$ is enough.
    \item Continuity only on $f$ is not enough. The counterexample in part (a) has $f$ continuous.
    
    Continuity on $g$ \textit{is} enough. If $g$ is continuous at $q$ then $\lim_{x \to q}g(x)=g(q)$.
    Then $\lim_{x \to q}g(x)=g(q)=\lim_{x \to p}g(f(x))$.
  \end{enumerate}
\end{solution}

%------------------------- Problem 5 -------------------------
\miquestion Show using Definition 4.3.1 that if $c$ is an isolated point of $A \subseteq \textbf{R}$ then $f: A \to \textbf{R}$ is continuous at $c$.
\begin{solution}
  This is vacuously true. By definition of isolated point, there's a neighborhood $V_{\delta}(c)$ such that the only point in it is $c$.

  Choose such a $\delta$, then $|x-c| < \delta \implies |f(x)-f(c)| < \epsilon$ is true because the only point in the interval is $x=c$ so
  $|f(x)-f(c)| = 0 < \epsilon$.
\end{solution}

%------------------------- Problem 6 -------------------------
\miquestion Provide an example of each or explain why the request is impossible.
\begin{enumerate}[label=(\alph*)]
  \item Two functions $f$ and $g$, neither of which is continuous at 0 but such that $f(x)g(x)$ and $f(x)+g(x)$ are continuous at 0.
  \item A function $f(x)$ continuous at 0 and $g(x)$ not continuous at 0 such that $f(x)+g(x)$ is continuous at 0.
  \item A function $f(x)$ continuous at 0 and $g(x)$ not continuous at 0 such that $f(x)g(x)$ is continuous at 0.
  \item A function $f(x)$ not continuous at 0 such that $f(x)+\frac{1}{f(x)}$ is continuous at 0.
  \item A function $f(x)$ not continuous at 0 such that $[f(x)]^{3}$ is continuous at 0.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item \[f(x) = 
        \begin{cases}
        0 & \text{if } x \ne 0 \\
        1 & \text{if } x = 0
        \end{cases}
      \]
      \[g(x) = 
        \begin{cases}
        1 & \text{if } x \ne 0 \\
        0 & \text{if } x = 0
        \end{cases}
      \]
    Both functions are not continuous at 0. But $f(x)+g(x)=1$ everywhere and $f(x)g(x)=0$ everywhere, both of which continuous everywhere (including at 0).
  
    \item Impossible. By the Theorem 4.3.4 (Algebraic Continuity Theorem), $-f(x)$ is continuous at 0 and $f(x)+g(x) -f(x) = g(x)$ is continuous at 0.
    \item $f(x)=0$ everywhere and $g(x)$ any function discontinuous at $0$ e.g. the example from part (a). Then $f(x)g(x)=0$ everywhere is continuous at 0.
    \item $f(x)=2$ for $x \geq 0$ and $f(x)=1/2$ otherwise. $f$ is discontinuous at 0, but $f(x)+\frac{1}{f(x)}=5/2$ everywhere.
    \item Impossible. Exercise 4.3.1 has proved that $g(x)=\sqrt[3]{x}$ is continuous everywhere. Then by Theorem 4.3.9 (Composition of Continuous Functions),
    if $[f(x)]^3$ continuous, $g(f(x)^3)=f(x)$ must be continuous as well.
  \end{enumerate}
\end{solution}

%------------------------- Problem 7 -------------------------
\miquestion
\begin{enumerate}[label=(\alph*)]
  \item Referring to the proper theorems, give a formal argument that Dirichlet's function from Section 4.1 is nowhere-continuous on \textbf{R}.
  \item Review the definition of Thomae's function in Section 4.1 and demonstrate that it fails to be continuous at every rational point.
  \item Use the characterization of continuity in Theorem 4.3.2 (iii) to show that Thomae's function is continuous at every irrational point in \textbf{R}. (given
  $\epsilon > 0$, consider the set of points $\{x \in \textbf{R} : t(x) \geq \epsilon\}.$)
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item It's discontinuous at all rational points. Pick $\epsilon = \frac{1}{2}$, then no $\delta$ works because for $x \in \textbf{Q}$, any neighborhood $V_{\delta}(x)$ must contain an
    an irrational point $y$ and $|f(x)-f(y)|=1 > \epsilon$.

    By a similar argument, it's also discontinuous at all irrational points.

    \item Consider a rational $\frac{p}{q}$ in its lowest terms, then $t(\frac{p}{q})=\frac{1}{q}$. Pick $\epsilon = \frac{1}{q+1}$, then no $\delta$ works because any neighborhood $V_{\delta}(\frac{p}{q})$
    must contain an an irrational point $y$ and $|f(\frac{p}{q})-t(y)|=\frac{1}{q} > \epsilon$.

    \item TODO
  \end{enumerate}
\end{solution}

%------------------------- Problem 8 -------------------------
\miquestion True or false with short proof or counterexample. Assume throughout that $g$ is defined on all of \textbf{R}
\begin{enumerate}[label=(\alph*)]
  \item If $g(x) \geq 0$ for all $x < 1$, then $g(1) \geq 0$ as well.
  \item If $g(r) = 0$ for all $r \in \textbf{Q}$, then $g(x)=0$ for all $x \in \textbf{R}$.
  \item If $g(x_{0}) > 0$ for a single point $x_{0} \in \textbf{R}$, then $g(x)$ is in fact strictly positive for uncountably many points.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item False. Let $g(x)=1$ for $x < 1$ and $g(x) = -1$ for $x \geq 1$. We have $g(x) \geq 0$ for all $x < 1$ but $g(1) = -1$.
    \item False. Let $g(r) = 0$ for all $r \in \textbf{Q}$ and $g(x) = 1$ for all irrational $x$.
    \item False. Let $g(x) = 0$ everywhere except $g(0)=1$.
  \end{enumerate}
\end{solution}

%------------------------- Problem 9 -------------------------
\miquestion Assume $h : \textbf{R} \to \textbf{R}$ is continuous on \textbf{R} and let $K = \{x : h(x)=0\}$. Show that $K$ is a closed set.
\begin{solution}
  Let $(x_{n})$ be a sequence in $K$ such that $(x_{n}) \to L$. Showing $K$ is closed amounts to showing $L \in K$.
  
  By continuity, $(x_{n}) \to L \implies h(x_{n}) \to h(L)$. Since $K$ is defined such that $h(x)=0$ for all $x \in K$, then $h(x_{n}) \to 0$ i.e. $h(L)=0$ so by definition $L \in K$.

  Note that there's nothing special about 0 here. The proof works just as well if we replace 0 with any other constant $c$.
\end{solution}

%------------------------- Problem 10 -------------------------
\miquestion Observe that if $a$ and $b$ are real numbers, then
\[\text{max}\{a, b\} = \frac{1}{2}[(a+b) + |a-b|]\]
\begin{enumerate}[label=(\alph*)]
  \item Show that if $f_{1}, f_{2}, ..., f_{n}$ are continuous functions, then
  \[g(x) = \text{max}\{f_{1}(x), f_{2}(x), f_{3}(x), ..., f_{n}(x)\}\]
  is a continuous function.
  \item Let's explore whether the result in (a) extends to the infinite case. For each $n \in \textbf{N}$, define $f_{n}$ on $\textbf{R}$ by
  \[
    f_{n}(x) = \begin{cases} 
    1 & \text{if } |x| \geq 1/n \\
    n|x| & \text{if } |x| < 1/n
    \end{cases}
  \]
  Now explicitly compute $h(x) = \text{sup}\{f_{1}(x), f_{2}(x), f_{3}(x), ...\}$.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item We first prove that $m(x) = \text{max}(a(x),b(x))$ is continuous for any continuous functions $a, b$. 
    
    --- INCORRECT --

    Once we've proved this, then we note that $g(x)$ is simply a composition of $m(x)$'s. Explicitly, define $m_{n}(x)=\text{max}(f_{n}(x), f_{n+1}(x))$. Then $g(x)=m_{n-1}(m_{n-2}(...(m_{1}(x))))$.
    By Theorem 4.3.9, composition of continous functions is continuous, so $g$ is continous.

    -- INCORRECT --

    -- CORRECTED -- 
  
    The above is not correct. For example, if $n=3$
    \[g(x) = m_{2}(m_{1}(x)) = \text{max}\{f_{2}(m_{1}(x)), f_{3}(m_{1}(x))\}\]
    whereas the correct one should be
    \[g(x) = m_{2}(x)=\text{max}\{m_{1}(x), f_{3}(x)\}\]

    What we need is recursion, not composition. Define $m_{1}(x)=\text{max}\{f_{1}(x), f_{2}(x)\}$ and
    \[m_{n}(x) = \text{max}\{m_{n-1}(x), f_{n+1}(x)\}\]
    Then $g(x)=m_{n}(x)$ is continuous by induction once we've proved that $m$ is continuous.

    -- CORRECTED -- 

    Now to prove continuity of $m$ at any $c \in \textbf{R}$. We need to prove that there exists $\delta$ such that $|x-c| < \delta \implies |\text{max}\{a(x), b(x)\} - \text{max}\{a(c), b(c)\}| < \epsilon$.
    
    Since $a$ and $b$ are continuous, there exists $\delta$ such that $a(x) \in V_{\epsilon}\{a(c)\}$ and $b_{x} \in V_{\epsilon}\{b(c)\}$. WLOG $b(c) \geq a(c)$.
    
    Since $b(x) < b(c) + \epsilon$, then $\text{max}\{a(x), b(x)\} < b(c) + \epsilon$

    Since $b(x) > b(c) - \epsilon$ and $\text{max}\{a(x), b(x)\} \geq b(x)$, then $\text{max}\{a(x), b(x)\} > b(c) - \epsilon$

    Combining them gives $b(c)-\epsilon < \text{max}\{a(x), b(x)\} < b(c)+\epsilon$. By assumption $\text{max}\{a(c), b(c)\} = b(c)$. Then $|\text{max}\{a(x), b(x)\} - \text{max}\{a(c), b(c)\}| < \epsilon$.
    
    Alternatively, use the given identity. By the Algebraic Continuity Theorem, $(a+b), (a-b)$ and $|a-b|$ are all continuous, so the $max$ function is continuous.

    --- Extra note ---

    I used the $\epsilon-\delta$ proof because I feel unsatisfied using the identity. Had it not been given, I wouldn't have come up with it. I wouldn't think to associate a $max$ function with an algebraic formula.

    But this formula actually has a simple geometric interpretation. To get the max between two numbers, start from their midpoint, then add half their distance. Likewise, to get the min, subtract half their distance.
    It wasn't immediately obvious because the way it's written in the problem statement, the $\frac{1}{2}$ was factored out.
  
    \item The point of this strange function is to show that part (a) actually does *not* extend to the infinite case.
    
    For $x=0$, $f_{n}(x)=0$ for all $n$, so $h(0)=0$.
    
    Now consider $x \neq 0$. There exists a rational $\frac{1}{n} \leq |x|$ for any $x \neq 0$. This means there's always an $n$ such that $f_{n}(x)=1$, which in turn means $h(x)=1$ for all $x \neq 0$.

    So $h(x)=1$ everywhere except $h(0)=0$, so $h$ is not continuous everywhere.

    Let's also see that the same argument doesn't work for the finite case. Suppose $h(x)=\text{max}\{f_{1}(x), f_{2}(x), ..., f_{n}(x)\}$. For $|x| \geq 1/n$, $h(x)=1$. For $x < 1/n$,
    we have $f_{n}(x)=n|x|$ the largest among $f_{1}(x), f_{2}(x), ..., f_{n}(x)$. So $h(x)$ is really equal $f_{n}(x)$, which is continuous everywhere.
  \end{enumerate}
\end{solution}


%------------------------- Problem 11 -------------------------
\miquestion TODO

%------------------------- Problem 12 -------------------------
\miquestion Let $F \subseteq \textbf{R}$ be a nonempty closed set and define $g(x) = inf\{|x-a| : a \in F\}$. Show that $g$ is continuous on all of \textbf{R} and
$g(x) \neq 0$ for all $x \notin F$.
\begin{solution}
  TODO
\end{solution}

%------------------------- Problem 14 -------------------------
\miquestion Let $f$ be a function defined on all of \textbf{R} that satisfies the additive condition $f(x+y) = f(x)+f(y)$ for all $x, y \in \textbf{R}$.
\begin{enumerate}[label=(\alph*)]
  \item Show that $f(0)=0$ and that $f(-x)=-f(x)$ for all $x \in \textbf{R}$.
  \item Let $k=f(1)$. Show that $f(n)=kn$ for all $n \in \textbf{N}$, and then prove that $f(z)=kz$ for all $z \in \textbf{Z}$. Now, prove that $f(r)=kr$ for any
  rational number $r$.
  \item Show that if $f$ is continuous at $x=0$, then $f$ is continuous at every point in \textbf{R} and conclude that $f(x)=kx$ for all $x \in \textbf{R}$. Thus,
  any additive function that is continuous at $x=0$ must necessarily be a linear function through the origin.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item The only solution to $f(0+0)=f(0)+f(0)$ is $f(0)=0$.
    
    $f(x-x)=f(0)=0=f(x)+f(-x)$. Let $f(x)=y$. Then only solution to $y+f(-x)=0$ is $f(-x)=-y=-f(x)$.

    \item $f(n) = \underbrace{f(1)+f(1)+ \cdots + f(1)}_{n \text{ times}} = nf(1)$.
    
    If $z$ is non-negative, then $z \in \textbf{N}$ and the proof is as above. Otherwise,
    \[f(z) = \underbrace{f(-1)+f(-1)+\cdots+f(-1)}_{|z| \text{ times}}=zf(1)\]
    Where we used the fact that $f(-1)=-f(1)$ from part (a).

    To prove $f(r) = rf(1)$ for any rational number $r$, first we prove that $f(zx)=zf(x)$ for any $z \in \textbf{Z}$ and $x \in \textbf{R}$. The proof is the same as before.
    For positive $z$:
    \[f(zx) = \underbrace{f(x)+f(x)+\cdots+f(x)}_{|z| \text{ times}}=zf(x)\]
    and for negative $z$, replace $f(x)$ with $f(-x)$.

    Now for any rational number $r=\frac{m}{n}$ and $m > 0$, we have
    \[f(\frac{m}{n}) = \underbrace{f(\frac{1}{n})+f(\frac{1}{n})+\cdots+f(\frac{1}{n})}_{m \text{ times}}=mf(\frac{1}{n})\]

    We still need to prove that $f(\frac{1}{n})=\frac{1}{n}f(1)$. How do we relate $\frac{1}{n}$ and $1$? Use the identity $n \cdot \frac{1}{n}=1$. We have
    $f(n \cdot \frac{1}{n})=f(1) \implies nf(\frac{1}{n})=f(1) \implies f(\frac{1}{n})=\frac{1}{n}f(1)$.

    Then $f(\frac{m}{n})=mf(\frac{1}{n})=\frac{m}{n}f(1)$, as required. The case for negative $m$ is very similar.

    *follow-up improvements:
    \begin{enumerate}
      \item For proving the $\textbf{Z}$ case for negative $z$, we can write $z=-m$ with $m \in \textbf{N}$, then $f(-m)=-f(m)=-mf(1)=zf(1)$. Slightly shorter.
      \item Likewise, to prove $f(zx)=zf(x)$ for negative $z$, use $z=-m$ again, then $f(zx)=-f(mx)=-mf(x)=zf(x)$.
      \item For the rational part, another way is to first prove that $nf(\frac{m}{n})=f(n \cdot \frac{m}{n})=f(m)=mf(1)$. Combining the first and last equality
      immediately gives the desired result.
    \end{enumerate}

    \item If $f$ is continuous at 0 then for any $\epsilon > 0$ there exists $\delta_{1} > 0$ such that $|x| < \delta_{1} \implies |f(x)| < \epsilon$.
    
    To prove $f$ is continuous at any $c \in \textbf{R}$, we need to prove for any $\epsilon > 0$ there exists $\delta_{2} > 0$ such that $|x-c| < \delta_{2} \implies |f(x)-f(c)| < \epsilon$.

    From the additive condition, $f(x)-f(c)=f(x-c)$. Set $y=x-c$ and $\delta_{2}=\delta_{1}$, then $|y| < \delta_{2} \implies |f(y)| < \epsilon$, as required.

    To prove $f(x)=f(1)x$ for all $x \in \textbf{R}$, we make use of the fact that for any real number $x$ there exists a sequence of rationals $(r_{n})$ that converges to $x$.
    Then $f(x) = \lim_{n \to \infty}f(r_{n})$ by continuity. Since $f(r_{n})$ is by definition $f(r_{1}), f(r_{2}), f(r_{3}), ...$, then
    \[f(x) = \lim_{n \to \infty} f(r_{1}), f(r_{2}), ... = \lim_{n \to \infty} r_{1}f(1), r_{2}f(1), ... = \lim_{n \to \infty} f(1)(r_{n})=f(1)x\]
  \end{enumerate}
\end{solution}

\end{questions}
\end{document}
