\documentclass{oxmathproblems}

\printanswers

\course{Understanding Analysis, 2nd Edition}
\sheettitle{Chapter 4 - Section 4.5 - The Intermediate Value Theorem} 

\begin{document}
\begin{questions}

%------------------------- Problem 1 -------------------------
\miquestion Show how the IVT follows as a corollary to Theorem 4.5.2.
\begin{solution}
  Theorem 4.5.2: continuous function preserves connected sets.

  Definition of connected sets: A set $E \subseteq \textbf{R}$ is connected iff whenever $a < c < b$ with $a, b \in E$, then $c \in E$ as well.

  In \textbf{R}, connected sets coincide precisely with the collection of intervals (page 104).
  
  Then IVT follows pretty much by definition. A closed interval $[a, b]$ is connected. Let $f$ be continuous, then $f([a, b])$ is connected. By definition of connectedness,
  $f$ contains everything in $[f(a), f(b)]$.
\end{solution}

%------------------------- Problem 2 -------------------------
\miquestion Example or prove impossible.
\begin{enumerate}[label=(\alph*)]
  \item A continuous function defined on an open interval with range equal to a closed interval.
  \item A continuous function defined on a closed interval with range equal to an open interval.
  \item A continuous function defined on an open interval with range equal to an unbounded closed set different from \textbf{R}.
  \item A continuous function defined on all of \textbf{R} with range equal to \textbf{Q}.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item $\sin(1/x)$ is continuous on $(0, 1)$ with range $[-1, 1]$.
    \item For unbounded interval, yes. $f(x)=x$ is continuous on \textbf{R} and its range is also \textbf{R} (recall that \textbf{R} is both open and closed).
    
    For unbounded interval, impossible. A closed bounded interval is compact and continuous functions preserve compactness (Theorem 4.4.1).
    \item With unbounded interval: $x^2$ is continuous on \textbf{R} with range $[0, \infty)$.
    
    With bounded interval: $\tan^2{x}$ is continuous on $(\frac{\pi}{2}, \frac{\pi}{2})$ with range $[0, \infty)$.

    \item Impossible. We'll prove it by contradiction. Suppose there exists such a function. Since the range is \textbf{Q}, $f$ is not constant. 
    Choose $x \neq y$ such that $f(x) \neq f(y)$ and wlog $f(x) < f(y)$.
    By IVT, the range of $f$ must contain everything in $[f(x), f(y)]$. But this implies $f$ contains an
    irrational, since between any two rationals there is an irrational, contradicting the assumption that the range is \textbf{Q}.
  \end{enumerate}
\end{solution}

%------------------------- Problem 3 -------------------------
\miquestion A function $f$ is \textit{increasing} on $A$ if $f(x) \leq f(y)$ for all $x < y$ in $A$. Show that if $f$ is increasing on $[a, b]$ and satisfies the
intermediate value property (Definition 4.5.3), then $f$ is continuous on $[a, b]$.
\begin{solution}
  To prove that $f$ is continuous at $a \leq c \leq b$, we need to show that given $\epsilon > 0$, there exists $\delta > 0$ such that $|x-c| < \delta \implies |f(x)-f(c)| < \epsilon$.

  We'll prove left continuity: there exists $\delta_{a}$ such that for $x < c$ and $c-x < \delta_{a}$, $f(c)-f(x) < \epsilon$. The proof for $x > c$ is similar and will result in a $\delta_{b}$. Then
  choose $\delta=\text{min}(\delta_{a}, \delta_{b})$.

  By monotonicity, $f(a) \leq f(c)$ and if $a < x < c$ then $f(a) \leq f(x) \leq f(c)$. By intermediate value property, $f$ takes on all values between $f(a)$ and $f(c)$ so there is an $a < x < c$
  such that $f(c)-f(x) < \epsilon$. Choose $\delta_{a} = c-x$.
  This works for all $x < t < c$ as well because monotonicity ensures that $f(c)-f(t) \leq f(c)-f(x) < \epsilon$.
\end{solution}

%------------------------- Problem 4 -------------------------
\miquestion Let $g$ be continuous on an interval $A$ and let $F$ be the set of points where $g$ fails to be one-to-one; that is,
\[F = \{x \in A : f(x)=f(y) \text{ for some } y \neq x \text{ and } y \in A\}.\]
Show $F$ is either empty or uncountable.

\begin{solution}
  At first I thought the problem statement must be wrong. Take something like $\sin(x)$. Within an interval there's only a finite number of points where $\sin(x)=1$, so it's neither empty nor uncountable. But then I realized
  it's not only 1 that's repeating, the intervals around $\sin(x)=1$ must be repeating as well and an interval is uncountable. That's what we'll prove; $F$ is either empty or contains repeating intervals of output values.

  If $g$ never fails to be one-to-one then $F$ is trivally empty. Otherwise there exists $x \neq y$ such that $g(x)=g(y)$.
  
  If for all $x < c < y$, $g(c)=g(x)=g(y)$, then $[x, y] \subseteq F$ so $F$ is uncountable.

  Otherwise, let $g(c) \neq g(x)$ and wlog $g(c) > g(x)$. By IVT there exists a subset $S_{1} \subseteq (x, c)$ such that $g(S_{1}) = (g(x), g(c))$. Since $g(S_{1})$ is an interval, $S_{1}$ must be uncountable.
  
  Similarly, there exists a subset $S_{2} \subseteq (c, y)$ such that $g(S_{2})=(g(y), g(c))$.
  
  Since $g(x)=g(y)$ then $g(S_{1})=g(S_{2})$, hence $S_{1}, S_{2} \subseteq F$ and $F$ is uncountable.  Note that $S_{1} \cap S_{2} = \emptyset$ so they're different set of points
  and can be included in $F$.
\end{solution}

%------------------------- Problem 5 -------------------------
\miquestion
\begin{enumerate}[label=(\alph*)]
  \item Finish the proof of the Intermediate Value Theorem using the Axiom of Completeness started previously.
  \item Finish the proof of the Intermediate Value Theorem using the Nested Interval Property started previously.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item The partial proof from the book is as follow. Let $f$ be continuous and consider the special case $f(a) < 0 < f(b)$. We'll show that $f(c)=0$ for some $c \in (a, b)$. Let
    \[K = \{x \in [a, b] : f(x) \leq 0\}.\]
    Notice that $K$ is bounded above by $b$ and $a \in K$ so $K$ is not empty. Thus we may appeal to AoC to assert that $s=\text{sup }K$ exists.

    We'll first prove that $f(s) \leq 0$. Since $s$ is a supremum, $s$ is a limit point of some sequence $(x_{n}) \in K$. By continuity, $f(x_{n}) \to f(s)$. Since
    $f(x_{n}) \leq 0$, then $f(s) \leq 0$.

    Next we need to prove that $f(s) \geq 0$. The idea is that since $s$ is a supremum, it's sort of the "rightmost" point where $f$ is $\leq 0$. But if $f(s) < 0$,
    since everything to the right of $s$ is $> 0$, there'd be a discontinuity; a jump that skips 0. Or in other words, $f(s)$ is not continuous from the right side.

    Formally, we prove by contradiction. First we note that since $s = \text{sup } K$, if $x > s$ then $x \notin K$ hence $f(x) > 0$.
    
    Now let $\epsilon = |f(s)|$. By continuity, there exists a $\delta < (b-s)$ that works (the bound $(b-s)$ is just a trivial detail to ensure
    we stay within $[a, b]$). Choose $x$ such that $x-s < \delta$. Since $f$ is defined on $[a, b]$, $f(x)$ is defined.
    
    We've asserted that $f(x) > 0$. But since $f(s) < 0$, we have $f(x)-f(s) \geq |0-f(s)| = \epsilon$, contradicting continuity of $f$.
    
    \item Repeat the procedure described in the book to get a sequence of nested intervals. Apply NIP. It's not difficult to conclude that 0 is the intersection.
  \end{enumerate}
\end{solution}

%------------------------- Problem 6 -------------------------
\miquestion Let $f: [0, 1] \to \textbf{R}$ be continuous with $f(0)=f(1)$.
\begin{enumerate}[label=(\alph*)]
  \item Show that there must exist $x, y \in [0, 1]$ satisfying $|x-y|=1/2$ and $f(x)=f(y)$.
  \item Show that for each $n \in \textbf{N}$ there exist $x_{n}, y_{n} \in [0, 1]$ with $|x_{n}-y_{n}|=1/n$ and $f(x_{n})=f(y_{n})$.
  \item If $h \in (0, 1/2)$ is not of the form $1/n$, there does not necessarily exist $|x-y|=h$ satisfying $f(x)=f(y)$. Provide an example that illustrates this using $h=2/5$.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item Use the same trick as problem 7 (which I had solved before this). Define $g(x)=f(x)-f(x+1/2)$ for $x \in [0, 1/2]$. Then the required condition is equivalent to the existence
    of $g(x)=0$.
    
    Consider $g(0)$ and $g(1/2)$. Since $f(0)=f(1)$, then $g(0)=-g(1/2)$ (both could possibly be 0 as well). Apply IVT to $[0, 1/2]$. Since $g(0)$ and $g(1/2)$
    have different signs, there's a $g(x)=0$.
  \end{enumerate}
\end{solution}

%------------------------- Problem 7 -------------------------
\miquestion Let $f$ be a continuous function on the closed interval $[0, 1]$ with range also contained in $[0, 1]$. Prove that $f$ must have a fixed point; that is, show $f(x)=x$ for
at least one value of $x \in [0, 1]$.
\begin{solution}
  Stuck with this for quite some time and finally got a hint from chatgpt: consider the auxiliary function $g(x)=f(x)-x$.

  With that hint, the problem becomes much easier. $f(x)=x$ if $g(x)=0$. Since the range of $f$ is in $[0, 1]$, then $0 \leq f(0), f(1) \leq 1$ and $g(1) \leq 0 \leq g(0)$.
  Note that $g$ is continuous, so we can apply IVT to $[0, 1]$ and conclude there exists an $0 \leq x \leq 1$ such that $g(x)=0$. 
\end{solution}

%------------------------- Problem 8 -------------------------
\miquestion \textbf{(Inverse functions).} If a function $f: A \to R$ is one-to-one (injective), the we can define the inverse function $f^{-1}$ on the range of $f$ in the
natural way: $f^{-1}(y)=x$ where $y=f(x)$. Show that if $f$ is continuous on an interval $[a, b]$ and one-to-one, then $f^{-1}$ is also continuous.

\begin{solution}
  Take a convergent sequence $(y_{n}) \to y$ and let $f^{-1}(y_{n})=(x_{n})$. To prove continuity of $f^{-1}$, we need to prove that $(y_{n}) \to y \implies (x_{n}) \to x$.

  \textbf{Wrong solution}\\
  My first stupid thought was to argue that since $f$ is continuous, then $(x_{n}) \to x \implies (f(x_{n}) = (y_{n})) \to (f(x)=y)$. But this argument is circular; we're using
  the thing we're trying to prove, mainly that $(x_{n}) \to x$. We *don't know* that $(x_{n}) \to x$.
  
  \textbf{Correct solution}\\
  In fact, we don't know yet that $(x_{n})$ converges at all. But we can use compactness: since $(x_{n}) \in [a, b]$ then there's a convergent subsequence $(x_{n_{k}}) \to L$.

  Our strategy will be to prove that $L=x$. Then we show that the whole sequence $(x_{n}) \to x$.

  By continuity of $f$, we have $(x_{n_{k}}) \to L \implies f(x_{n_{k}}) \to f(L)$. But note that since $f(x_{n_{k}})$ is really a subsequence of $(y_{n})$ and
  by assumption $(y_{n}) \to y$, then $f(x_{n_{k}}) \to f(L) = y = f(x)$, hence $L=x$.

  Note that the above argument works for any arbitrary convergent subsequence $(x_{n_{k}}) \in (x_{n})$. Finally, we invoke Exercise 2.5.5 (which I haven't solved!): if a
  bounded sequence $(x_{n})$ has the property that every convergent subsequence converges to the same limit $L \in \textbf{R}$, then $(x_{n}) \to L$. Here $(x_{n})$ is bounded by
  $[a, b]$ and $L=x$.
\end{solution}

\end{questions}
\end{document}
