\documentclass{oxmathproblems}

\printanswers

\course{Understanding Analysis, 2nd Edition}
\sheettitle{Chapter 5 - Section 5.2 - Derivatives and the Intermediate Value Property} 

\begin{document}
\begin{questions}

%------------------------- Problem 1 -------------------------
\miquestion Supply proofs for part (i) and (ii) of Theorem 5.2.4.
\begin{solution}
  Part (i): $(f+g)'(c) = f'(c) + g'(c)$

  Proof:
  \begin{align*}
     (f+g)'(c) = \lim_{x \to c}\frac{f(x)+g(x)-f(c)-g(c)}{x-c} &= \lim_{x \to c}\left(\frac{f(x)-f(c)}{x-c} + \frac{g(x)-g(c)}{x-c}\right)\\ 
     &= f'(c)+g'(c)
  \end{align*}
  where the last step is justified because we know both limits exist, so we can apply the Algebraic Limit Theorem (Theorem 4.2.4).

  Part (ii): $(kf)'(c) = kf'(c)$ for all $k \in \textbf{R}$.
  
  Proof:
  \[(kf)'(c) = \lim_{x \to c} k\left( \frac{f(x)-f(c)}{x-c} \right) = k \lim_{x \to c}\frac{f(x)-f(c)}{x-c} = kf'(c)\]
\end{solution}

%------------------------- Problem 2 -------------------------
\miquestion Exactly one of the following requests is impossible. Decide which it is, and provide examples for the other three. In each case, assume the functions are defined
on all of \textbf{R}.
\begin{enumerate}[label=(\alph*)]
  \item Functions $f$ and $g$ not differentiable at zero but where $fg$ is differentiable at zero.
  \item A function $f$ not differentiable at zero and a function $g$ differentiable at zero where $fg$ is differentiable at zero.
  \item A function $f$ not differentiable at zero and a function $g$ differentiable at zero where $f+g$ is differentiable at zero.
  \item A function $f$ differentiable at zero but not differentiable at any other point.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item $f(x) = 0$ for $x \geq 0$ and $f(x)=1$ otherwise. $g(x)=1$ for $x \geq 0$ and $g(x) = 0$ otherwise. Both functions are discontinuous at zero hence not differentiable there, but
    $fg$ is the zero function.
    \item $f(x) = 0$ for $x \geq 0$ and $f(x)=1$ otherwise. $g(x)=0$ everywhere.
    \item Impossible. If $f+g$ were differentiable at zero, then $(f+g)-g = f$ must be differentiable at zero as well by Theorem 5.2.4.
    \item $f(x) = x^2$ for $x \in \textbf{Q}$ and $f(x) = 0$ otherwise.
    
    It's differentiable at zero because for $c= 0$, $\frac{f(x)-f(c)}{x-c} = x$ and $\lim_{x \to 0} x = 0$.

    It's not differentiable at $c \neq 0$. Consider a sequence $(x_{n}) \to c$. We can split it into two subsequences $(a_{n}) \in \textbf{Q}$ and $(b_{n}) \notin \textbf{Q}$. We have
    $(a_{n}) \to c^2$ and $(b_{n}) \to 0$. Since $c^2 \neq 0$ for $c \neq 0$, the two subsequences converge to different limits, hence $(x_{n})$ doesn't converge.
  \end{enumerate}
\end{solution}

%------------------------- Problem 3 -------------------------
\miquestion
\begin{enumerate}[label=(\alph*)]
  \item Use Definition 5.2.1 to produce the proper formula for the derivation of $f(x)=1/x$.
  \item Combine the result in part (a) with the Chain Rule to supply a proof for part (iv) of Theorem 5.2.4.
  \item Supply a direct proof of Theorem 5.2.4 (iv) by algebraically manipulating the difference quotient for $(f/g)$ in a style similar to the proof of Theorem 5.2.4 (iii).
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item \[\frac{f(x)-f(c)}{x-c}=\frac{\frac{1}{x}-\frac{1}{c}}{x-c}=\frac{-(x-c)}{xc}\frac{1}{x-c}=-\frac{1}{xc}\]

    So we have $f'(c)=\lim_{x \to c}(-\frac{1}{xc})=-\frac{1}{c^2}$ for $c \neq 0$.

    \item We first derive a formula for the derivative of $h(x)=1/g(x)$.
    
    \[\frac{h(x)-h(c)}{x-c}=\frac{-(g(x)-g(c))}{g(x)g(c)}\frac{1}{x-c}=-\frac{1}{g(x)g(c)} \cdot \frac{g(x)-g(c)}{x-c}\]

    Taking the limit to $c$ and noting that the second term of the product is the definition of $g'(c)$
    \[h'(c) = \lim_{x \to c} -\frac{1}{g(x)g(c)} \cdot \frac{g(x)-g(c)}{x-c} = -\frac{g'(c)}{g(c)^2}\]

    Now apply the product rule to $(fh)'(c)$ to get the formula for $(f/g)'(c)$.

    ----- Simpler solution by chatgpt -----\\
    Derive the formula of $h'(x)$ by defining $h(x)=1/x$ then apply the chain rule to $h(g(x))$.

    \item
    \begin{align*}
       \frac{\frac{f}{g}(x) - \frac{f}{g}(c)}{x-c} &=  \frac{\frac{f}{g}(x) - \frac{f(x)}{g(c)} + \frac{f(x)}{g(c)} - \frac{f}{g}(c)}{x-c}\\
       &= \frac{f(x)}{x-c}\left(\frac{g(c)-g(x)}{g(x)g(c)}\right) + \frac{1}{g(c)(x-c)}\left(f(x)-f(c)\right)\\
       &= -\frac{f(x)}{g(x)g(c)}\left( \frac{g(x)-g(c)}{x-c} \right) + \frac{1}{g(c)} \left( \frac{f(x)-f(c)}{x-c} \right)
    \end{align*}

    Take the limit to $c$ to get
    \begin{align*}
      -\frac{f(c)g'(c)}{g(c)^2} + \frac{f'(c)}{g(c)} = \frac{f'(c)g(c)-f(c)g'(c)}{g(c)^2}
    \end{align*} 
  \end{enumerate}
\end{solution}

%------------------------- Problem 4 -------------------------
\miquestion Follow these steps to provide a slightly modified proof of the Chain Rule.
\begin{enumerate}[label=(\alph*)]
  \item Show that a function $h: A \to \textbf{R}$ is differentiable at $a \in A$ if and only if there exists a function $l : A \to \textbf{R}$ which is continuous at $a$ and satisfies
  \[h(x)-h(a) = l(x)(x-a) \qquad \text{for all } x \in A.\]
  \item Use this criterion for differentiability (in both directions) to prove Theorem 5.2.5.
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item First we prove that if such a function $l$ exists then $h$ is differentiable at $a$. We have
    \[h'(a) = \lim_{x \to a}\frac{h(x)-h(a)}{x-a} = \lim_{x \to a}l(x) = l(a)\]

    Next, we prove that if $h$ is differentiable at $a$ then such a function $l$ exists. Define
    \[
      l(x) = 
      \begin{cases}
        \frac{h(x)-h(a)}{x-a} & x \neq a\\
        h'(a) & (x=a)
      \end{cases}
    \]

    $l$ is continuous at $a$ since $\lim_{x \to a}l(x) = \lim_{x \to a}\frac{h(x)-h(a)}{x-a} = h'(a) = l(a)$.

    $h(x)-h(a)=l(x)(x-a)$ for all $x \in A$ by definition.

    \item TODO.
  \end{enumerate}
\end{solution}

%------------------------- Problem 5 -------------------------
\miquestion Let $ l(x) = \begin{cases} x^a & x > 0\\0 & x \leq 0 \end{cases}$.
\begin{enumerate}[label=(\alph*)]
  \item For which values of $a$ is $f$ continuous at zero?
  \item For which values of $a$ is $f$ differentiable at zero? In this case, is the derivative function continuous?
  \item For which values of $a$ is $f$ twice-differentiable?
\end{enumerate}

\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item $a > 0$. The limit at 0 from the left is 0. If $a < 0$ then the right limit diverges. If $a=0$ then the right limit is 1. But if $a > 0$ then the right limit is 0.
    \item From the direction of positive $x$, we have $\lim_{x \to 0^{+}}\frac{x^a-0^a}{x-0}=\lim_{x \to 0^{+}}x^{a-1} = 0^{a-1}$. This is only defined for $a > 1$, in which case it's 0 and
    agrees with the left limit, so the derivative is defined.

    The derivative is continuous, but seems to require us to use the fact that $l'(x)=ax^{a-1}$ for \textit{non-integer} $a$, which we have not proved yet.
    \item The next right-side derivative at zero is $\lim_{x \to 0^{+}} x^{a-2} = 0^{a-2}$, so we need $a > 2$.
  \end{enumerate}
\end{solution}

\end{questions}
\end{document}
